Refining from the BON High-level Specification Language to the VHDL-AMS Hardware Description Language

### 2009-08-26 papers/BON_VHDL/paper.tex
intro elaborated

%======================================================================
\section{Introduction}
The \ST sensor platform, represents heterogeneous collection of digital, analog 
and mixed signal hardware circuits. Embedded system complexity 
requires the need to strive for formal descriptions each component 
separatelly in order to produce specification of the \STS. 
Our aim is to be able to precisely describe the structures of circuits,
the behaviour of circuits and the formal link between structures and behaviours.
In other words, the behavioural specification of the \STS may be 
specified as a \inv on the signals at the circuit's ports and may be obtained 
by taking the conjunction of the \emph{invariants} that describe 
behaviours of the individual components separately. 
For this purpose we need  sophisticated tool and methodology 
in order to be able to achieve this goal.  
The authors will illustrate in this article how BON can represent a high 
level analogue mapping and will further illustrate a procedure for 
transformation from VHDL-AMS to BON.\\ 
This article provides two main contributions.
\begin{enumerate}
\item 
 It provides a solution how to formally emulate
 piece of hardware using BON-JML approach; up to date BON and JML
 has been used mostly in the pure object oriented software construction.    
\item
It relates hardware description code and formal calculus. 
This means that if there  is a code that is "correct by construction", 
then it will be derivable from informal BON specification, 
via the refinement calculus to formal BON specification 
and JML implementation.     
\end{enumerate}
As an extremly simple example, 
we will give practical aspect of refinement, applied on real capacitor component, 
in order to produce formal specification  and software simulation.\\  
A comprehensive list of the mapping between BON and VHDL-AMS can be 
found in the appendix.
%=====================================================================

\section{VHDL-AMS}
IEEE VHDL-AMS 1076.1 is the industry standard mixed-signal 
high-level description language for electronic and multi-domain 
systems. VHDL-AMS, as an superset of VHDL, supports modeling and 
simulation of analog and mixed-signal systems. The main contribution 
of VHDL-AMS is ability to describe the behaviour of complex system 
trought differential and algebraic equations (DAEs). For this purpose, 
VHDL-AMS introduces the following capabilities and therms:

@@

\section{Choice of BON as the modeling tool}
%Brief description and motivation.
The Business Object Notation (BON) is an analysis and design notation 
for object-oriented systems  which emphasizes \emph{seamlessness},\xspace 
\emph{reversibility} and software \emph{contracting}. 
Let us define these principles precisely and point out why the principles 
are important in VHDL-AMS simulation.

\subsection{Seamlessness and Reversibility}
\emph{Seamlessness} is the use of a continuous process 
throughout the software lifecycle.
\emph{Reversibility} is the support for both forward and 
backward development process: from analysis to design 
and implementation, and back.
In VHDL-AMS, modelling represents an continious top-down-bottom-up
process. Top-down approach serves for design while bottom-up
serves for verification. The possibility of seamlessly translating 
changes made during a certain development phase back into earlier phases,
facilitates a stable and predictable environment and ensures the various 
models consistent with the code. Keeping in sync various models, documetation
and specification from forward progress to backward and redesign, preclude 
that design and implementation diverge to an inconsistent state. 
\emph{Seamlessness} also plays an important role in relating behaviours at
different levels of abstraction. For example, one may wish to analyze
a "real capacitor" at the level of individual components (a resistor and ideal capacitor)
and their behaviours, or at the more abstract level through general behaviour (\inv) and
using existential quantification (\texttt{exists}) to encapsulate all internal signals. 
Therefore capturing similarities between the differing levels of abstractions 
far outweigh their inevitable differences, making it possible to obtain a 
continuous process that ensures a direct mapping between a formal specification 
and its software solution.

\subsection{Contracting}
\emph{Contracting} is the definition, for each software element, 
of the associated semantic properties. BON use a full assertion mechanism, 
allowing analysts to specify not only the structure of a system but also its
semantics ( constraints and invariants of the expected results ). 
In VHDL-AMS \emph{simultaneous statement} represents analogue constraint of 
continuous-time waveform. In other words, analog kernel calculate quantities such 
that the constraints specified by the simultaneous statements are satisfied. 
We will agree that the constraint of waveform is nothing more than \inv.
Another example of \inv is conservation of energy law in which 
the summation of \arcoss quantities (\texttt{$\sum_{loop}$ across quatities =0}) 
around any loop equals zero and the summation of through 
quantities (\texttt{$\sum_{node}$ through quatities =0}) at any node 
(\texttt{terminal}) equals zero. Moreover, any predefined (implicite) attribute in VHDL-AMS  
has defined pre (post) condition trought input (output) 
parameters and theirs constraints. Furthermore any subtype (subnature) 
represents restricted set of values from a base type (nature). 
The condition that determines which values are in the subtype (subnature)  
may be contracted by specifying characteristic predicate which represents \inv.\\ 
Another advantage of BON is an easily-readable, English-based textual notation. 
The current IEEE standard VHDL-AMS language reference manual tries to define 
VHDL-AMS as well as possible in a descriptive way, explaining the 
semantics in English.\\
Therefore, on the basis of the above, BON is the obvious choice 
for the \STS.
Throughput this article we will make this more clearly.  
\section{Mapping between VHDL-AMS and BON}
The main theme of this article is the mapping between VHDL-AMS and BON.
\section{Refinement from VHDL-AMS to BON }

An VHDL-AMS design units are described using an \ent-\arch structure. 
The design unit contains an \ent and one or more \archs. 
The \ent specifies the interface that an entity provides of itself 
to the outside. In other words, \ent introduces a name, list of 
input/output ports and the definition of the generic parameters. 
The textual presentation of \ent is shown in Figure.1. 
\input{adc-entity}

>>>>

%======================================================================
\section{Introduction}
\label{sec:introduction}
The \ST sensor platform, represents heterogeneous collection of digital, analog 
and mixed signal hardware circuits. Embedded system complexity 
requires the need to strive for formal descriptions each component 
separatelly in order to produce specification of the \STS. 
Our aim is to be able to precisely describe the structures of circuits,
the behaviour of circuits and the formal link between structures and behaviours.
In other words, the behavioural specification of the \STS may be 
specified as a \inv on the signals at the circuit's ports and may be obtained 
by taking the conjunction of the \emph{invariants} that describe 
behaviours of the individual components separately. 
For this purpose we need  sophisticated tool and methodology 
in order to be able to achieve this goal.  
The authors will illustrate in this article how BON can represent a high 
level analogue mapping and will further illustrate a procedure for 
transformation from VHDL-AMS to BON.\\ 
This article provides two main contributions.
\begin{enumerate}
\item 
 It provides a solution how to formally emulate
 piece of hardware using BON-JML approach; up to date BON and JML
 has been used mostly in the pure object oriented software construction.    
\item
It relates hardware description code and formal calculus. 
This means that if there  is a code that is "correct by construction", 
then it will be derivable from BON (informal/formal) specification, 
via the refinement calculus to JML implementation.     
\end{enumerate}
The rest of this article is organized as follows.
In \autoref{sec:vhdlams} we have  presented an overview of VHDL-AMS language.
\autoref{sec:choiceobon} defines  BON principles and explains the 
reasons why we have decided to use BON method.
The central part of the our article is \autoref{sec:mappingVB}, which 
shows mapping between VHDL-AMS and BON. 
In \autoref{sec:refVB} we will give practical aspect of refinement, 
applied on real capacitor component, in order to produce formal specification  
and software simulation.\\  
A comprehensive list of the mapping between BON and VHDL-AMS can be 
found in the appendix.
%=====================================================================

\section{VHDL-AMS}
\label{sec:vhdlams}
IEEE VHDL-AMS 1076.1 is the industry standard mixed-signal 
high-level description language for electronic and multi-domain 
systems. VHDL-AMS, as an superset of VHDL, supports modeling and 
simulation of analog and mixed-signal systems. The main contribution 
of VHDL-AMS is ability to describe the behaviour of complex system 
trought differential and algebraic equations (DAEs). For this purpose, 
VHDL-AMS introduces the following capabilities and therms:

@@

\section{Choice of BON as the modeling tool}
\label{sec:choiceobon}
%Brief description and motivation.
The Business Object Notation (BON) is an analysis and design notation 
for object-oriented systems  which emphasizes \emph{seamlessness},\xspace 
\emph{reversibility} and software \emph{contracting}. 
Let us define these principles precisely and point out why the principles 
are important in VHDL-AMS simulation.

\subsection{Seamlessness and Reversibility}
\label{subsec:seamrever}
\emph{Seamlessness} is the use of a continuous process 
throughout the software lifecycle.
\emph{Reversibility} is the support for both forward and 
backward development process: from analysis to design 
and implementation, and back.
In VHDL-AMS, modelling represents an continious top-down-bottom-up
process. Top-down approach serves for design while bottom-up
serves for verification. The possibility of seamlessly translating 
changes made during a certain development phase back into earlier phases,
facilitates a stable and predictable environment and ensures the various 
models consistent with the code. Keeping in sync various models, documetation
and specification from forward progress to backward and redesign, preclude 
that design and implementation diverge to an inconsistent state. 
\emph{Seamlessness} also plays an important role in relating behaviours at
different levels of abstraction. For example, one may wish to analyze
a "real capacitor" at the level of individual components (a resistor and ideal capacitor)
and their behaviours, or at the more abstract level through general behaviour (\inv) and
using existential quantification (\texttt{exists}) to encapsulate all internal signals. 
Therefore capturing similarities between the differing levels of abstractions 
far outweigh their inevitable differences, making it possible to obtain a 
continuous process that ensures a direct mapping between a formal specification 
and its software solution.

\subsection{Contracting}
\label{subsec:contracting}
\emph{Contracting} is the definition, for each software element, 
of the associated semantic properties. BON use a full assertion mechanism, 
allowing analysts to specify not only the structure of a system but also its
semantics ( constraints and invariants of the expected results ). 
In VHDL-AMS \emph{simultaneous statement} represents analogue constraint of 
continuous-time waveform. In other words, analog kernel calculate quantities such 
that the constraints specified by the simultaneous statements are satisfied. 
We will agree that the constraint of waveform is nothing more than \inv.
Another example of \inv is conservation of energy law in which 
the summation of \arcoss quantities (\texttt{$\sum_{loop}$ across quatities =0}) 
around any loop equals zero and the summation of through 
quantities (\texttt{$\sum_{node}$ through quatities =0}) at any node 
(\texttt{terminal}) equals zero. Moreover, any predefined (implicite) attribute in VHDL-AMS  
has defined pre (post) condition trought input (output) 
parameters and theirs constraints. Furthermore any subtype (subnature) 
represents restricted set of values from a base type (nature). 
The condition that determines which values are in the subtype (subnature)  
may be contracted by specifying characteristic predicate which represents \inv.\\ 
Another advantage of BON is an easily-readable, English-based textual notation. 
The current IEEE standard VHDL-AMS language reference manual tries to define 
VHDL-AMS as well as possible in a descriptive way, explaining the 
semantics in English.\\
Therefore, on the basis of the above, BON is the obvious choice 
for the \STS.
Throughput this article we will make this more clearly.  
\section{Mapping between VHDL-AMS and BON}
\label{sec:mappingVB}
The main theme of this article is the mapping between VHDL-AMS and BON.
\section{Refinement from VHDL-AMS to BON }
\label{sec:refVB} 
An VHDL-AMS design units are described using an \ent-\arch structure. 
The design unit contains an \ent and one or more \archs. 
The \ent specifies the interface that an entity provides of itself 
to the outside. In other words, \ent introduces a name, list of 
input/output ports and the definition of the generic parameters. 
The textual presentation of \ent is shown in Figure.1. 
\input{adc-entity}
